---
title: "Mapping NZ Roads with sf"
editor: visual
author: "Synergia"
format: 
  html:
    embed-resources: true
    date: last-modified
    date-format: "DD-MM-YYYY"
    toc: true  
    toc_float:
      collapsed: true
      smooth_scroll: false
    css: 
      - "G:/Shared drives/Assets Library/Design/quarto_css/synergia_style_revised.css"
      - "G:/Shared drives/Assets Library/Design/quarto_css/custom.scss"
---

```{R}
pacman::p_load(tidyverse,here,glue,ggtext,showtext,extrafont,sf,ggplot2,ggfx,camcorder)
tt_source <- "Paul Beere, PhD. GeoHealth Laboratory"
bg_color <- adjustcolor("black",alpha.f = 1)
txt_color <- "#FF5674"
txt_color2 <- "white"

yellow <- "#FFB702"  
blue <- "#1E9BD5"
green <- "#20C29D"
light_blue <- "#00E7EC"
red <- "#FF5674"

tt_family <- "montserrat"
tt_family_2 <-  "roboto"


font_add_google("Montserrat", tt_family)
font_add_google("Roboto", tt_family_2)

font_add("fa-brands",
         regular = here("Assets","Fonts","fontawesome-free-6.4.2-web","webfonts","fa-brands-400.ttf"))
font_add("fa-solid",
         regular = here("Assets","Fonts","fontawesome-free-6.4.2-web","webfonts","fa-solid-900.ttf"))
font_add("fa-regular",
         regular = here("Assets","Fonts","fontawesome-free-6.4.2-web","webfonts","fa-regular-400.ttf"))
showtext_auto(enable = T) 

X  <- str_glue("<span style='font-family:fa-brands'>&#xe61b;</span>")
GH <- str_glue("<span style='font-family:fa-brands'>&#xf09b;</span>")
MN <- str_glue("<span style='font-family:fa-brands'>&#xf4f6;</span>")
DT <- str_glue("<span style='font-family:fa-brands'>&#xf1c0;</span>")

tt_caption <- str_glue(
  "Steven Turnbull #MappingNZ - 2023-11-06<br>
  {X} @Turdbull {GH} @sturbull {MN} @turdbull<br>
  Data: Beere, P., 2016: <br>
  Creating a Road Network Analysis Layer with <br>
  Travel Time Estimates using Open-source Data.<br>
  GeoHealth Laboratory, New Zealand
  "
)

tt_title_text    <- str_glue("Roads of Aotearoa New Zealand")

tt_subtitle_text <- str_glue(
  "Including <span style='color:#FFC300;'>Highway</span>, <span style='color:#FF5733;'>Principal</span>, <span style='color:#C70039;'>Arterial</span>, <span style='color:#900C3F;'>Collector</span>, and <span style='color:#581845;'>Residential</span> Roads."
)

```


## Introduction to `sf`

The `sf` package, which stands for Simple Features, is an R package that standardises how spatial data is handled and facilitates easy manipulation and visualisation of geographic data.

## Packages

To use the `sf` package for spatial data analysis and plotting maps, you should first install it from CRAN.

## Reading a Shapefile

To work with shapefiles, `sf` provides a function to read them directly into R as `sf` objects. Here's how to read the `nz_roads.shp` file:

```{r}


nz_roads <- st_read("G:/Shared drives/Assets Library/Geospatial/ShapeFiles/NZ_roads.shp")
```


## Plotting the Shapefile

Once the data is read into R, plotting is straightforward using the `plot` method from `sf` or integrating with `ggplot2`:

```{R}
gg_record(dir = here("outputs","2023-11-06"),
          device = "png",
          width = 5.75, height = 10, 
          units = "in", dpi = 900)
```

```{r}
# Or using ggplot2 for more control and aesthetics

nz_roads_clean <- nz_roads %>%
  mutate(road_group = case_when(
    str_detect(road_class,"Major") ~ "Highway",
    str_detect(road_class,"Principal") ~ "Principal",
    str_detect(road_class,"Arterial") ~ "Arterial",
    str_detect(road_class,"Collector") ~ "Collector",
    str_detect(road_class,"Residentia") ~ "Residential"
    )
    ) %>%
  
  mutate(road_size = case_when(
    str_detect(road_group,"Highway") ~ 0.7,
    str_detect(road_group,"Principal") ~ 0.7,
    str_detect(road_group,"Arterial") ~ 0.3,
    str_detect(road_group,"Collector") ~ 0.3,
    str_detect(road_group,"Residentia") ~ 0.1
    )
    ) %>%
   mutate(road_colour = case_when(
    str_detect(road_group,"Highway") ~ "#FFC300",
    str_detect(road_group,"Principal") ~ "#FF5733",
    str_detect(road_group,"Arterial") ~ "#C70039",
    str_detect(road_group,"Collector") ~ "#900C3F",
    str_detect(road_group,"Residentia") ~ "#581845"
    )
    )


ggplot() +
  geom_sf(data = nz_roads_clean,
          aes(color = road_colour,linewidth=road_size)) + # Replace with the actual variable of interest
  theme_void() +
  scale_colour_identity() +
  scale_size_identity() +
  scale_linewidth_identity() +
  labs(title = tt_title_text, 
       subtitle = tt_subtitle_text,
       caption = tt_caption) +
  #set theme
  theme(
    plot.background = element_rect(colour=bg_color,fill = bg_color),
    panel.background = element_rect(colour=bg_color,fill = bg_color),
    
    plot.title.position = "plot",
    plot.title = element_textbox_simple(
      halign = 0.5,
      tt_family,
      lineheight = 0.2,
      color = txt_color2,
      face = "bold",
      size = 220,
      margin = margin(t = 3, r = 0, b = 0, l = 0)
    ),
    
    plot.subtitle = element_textbox_simple(
      halign = 0.5,
      tt_family,
      color = txt_color2,
      #face = "bold",
      size = 180,
      lineheight = 0.1,
      margin = margin(t = 2, r = 0, b = 0, l = 0)
    ),
    
    plot.caption = element_textbox_simple(
      family = tt_family, 
      color = txt_color,
      size = 50,
      lineheight = 0.2,
      linewidth = 5,
      margin = margin(t = 1, r = 0, b = 3, l = 0)
    ),
    plot.caption.position = "plot",
    
    legend.position = "none"
    
  )
```

```{R}
camcorder::gg_playback()

```
